Optimal management of a stochastically varying population when policy adjustment is costly

### 2014-06-04 manuscripts/manuscript.Rmd
source code externally (easier for collaborators to edit)

output:
  pdf_document:
    keep_tex: true
    template: components/elsarticle.latex
---

<!-- CODE -->

```{r dependencies, include=FALSE}
#source("components/install.R")
```

```{r caching, include=FALSE}
library("methods")
library("knitr")
basename <- "manuscript"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)

# PDF-based figures
opts_chunk$set(dev='pdf')
```

```{r libraries, include=FALSE, cache=FALSE}
rm(list=ls())
library("rmarkdown")
library("pdgControl")
library("reshape2")
library("plyr")
library("ggplot2")
library("data.table")
library("pander")
library("cboettigR")
library("ggthemes")
library("snowfall")
theme_set(theme_tufte())
ncpu = 24
```

``` {r profit_model}
price = 10
c0 = 30
profit <- profit_harvest(price = price, c0 = c0, c1 = 0)
```

```{r c2_grid}
c2 <- exp(seq(0, log(41), length.out = 40))-1
c2 <- seq(0, 40, length.out=100)
```

```{r reduction}
reduction <- 0.25 
```

```{r setup}
seed <- 123                 # Random seed (replicable results)
delta <- 0.05               # economic discounting rate
OptTime <- 20               # stopping time
gridsize <- 50              # grid size for fish stock and harvest rate (discretized population)
sigma_g <- 0.2              # Noise in population growth
reward <- 0                 # bonus for satisfying the boundary condition
z_g <- function() rlnorm(1,  0, sigma_g) # mean 1
z_m <- function() 1         # No measurement noise, 
z_i <- function() 1         # No implemenation noise
f <- BevHolt                # Select the state equation
pars <- c(1.5, 0.05)        # parameters for the state equation
K <- (pars[1] - 1)/pars[2]  # Carrying capacity (for reference 
xT <- 0                     # boundary conditions
x0 <- K
x_grid <- seq(0.01, 1.2 * K, length = gridsize)  
h_grid <- seq(0.01, 0.8 * K, length = gridsize)  
````

```{r reed, dependson=c("setup", "profit_model")}
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g )
opt <- find_dp_optim(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                     profit, delta, reward=reward)
````

```{r fees}
L1 <- function(c2) function(h, h_prev)  c2 * abs(h - h_prev) 
fixed <-  function(c2) function(h, h_prev) c2 * as.numeric( !(h == h_prev) )
L2 <- function(c2) function(h, h_prev)  c2 * (h - h_prev) ^ 2
none <- function(h, h_prev)  0
penaltyfns <- list(L2=L2, L1=L1, fixed=fixed)
````

```{r parallel, include=FALSE}
sfInit(cpu=ncpu, parallel=T)
sfLibrary(pdgControl)
sfExportAll()
````

``` {r bigloop, dependson=c("setup", "reed", "fees", "profit_model", "c2_grid")}
policies <- lapply(penaltyfns, function(penalty){
  sfLapply(c2, function(c2){
      policy <- optim_policy(SDP_Mat, x_grid, h_grid, OptTime, xT, 
                   profit, delta, reward, penalty = penalty(c2))
      }
  )
})
````

```{r, dependson="quadcosts"}
i <- which(x_grid > K)[1]
fees <- 
lapply(policies, function(penalty) 
  sapply(penalty, function(c2_run)
    max(c2_run$V[i,]) # Would be penalty_free_V originally   ## this isn't correct for asym cases
  )
)
```

``` {r npv-plot, dependson="quadcosts"}
npv0 <- max(fees$L1) # all have same max, at c2=0 
fees <- data.frame(c2=c2,fees)
fees <- melt(fees, id="c2")
fees <- subset(fees, variable %in% c("L1", "L2", "fixed"))

#ggplot(fees, aes(c2, value, col=variable)) + geom_point() + geom_line()
````

```{r apples, dependson=c("quadcosts", "reduction")}
closest <- function(x, v){
  which.min(abs(v-x))
}
dt_npv <- data.table(fees)
index <- dt_npv[,closest(reduction, (npv0-value)/npv0), by=variable]
apples_index <- index$V1
names(apples_index) <- index$variable
apples <- c2[index$V1]
names(apples) <- index$variable
````

```{r print_npv, dependson="apples"}
setkey(dt_npv, variable)
values <- apply(index, 1, function(x) dt_npv[x[1], ][as.integer(x[2]), ]$value)
percent.error <- (values - ((1-reduction)*npv0)) / ((1-reduction)*npv0)* 100
print_npv <- data.frame(model=index$variable, "value realized"=values, 
                        "percent of npv0" = 100*values/npv0, "percent error"=percent.error)

```{r npv_table, dependson="print_npv", results='asis'} 
#pandoc.table(print_npv)
```

```{r policynames}
L2_policy <- policies$L2[[apples_index["L2"]]]$D
L1_policy <- policies$L1[[apples_index["L1"]]]$D
fixed_policy <- policies$fixed[[apples_index["fixed"]]]$D
```

```{r simulate_policy, dependson=c("policynames", "apples")}
reps <- 1:100
names(reps) = paste("rep", 1:length(reps), sep="_") # treat as a factor
seeds <- 1:100
sims <- list(
  L1 = lapply(reps, function(x) simulate_optim(f, pars, x_grid, h_grid, x0, 
                      L1_policy, z_g, z_m, z_i, 
                      opt$D, profit=profit, penalty=L1(apples["L1"]), seed=seeds[x])), 
  L2 = lapply(reps, function(x) simulate_optim(f, pars, x_grid, h_grid, x0, 
                      L2_policy, z_g, z_m, z_i, 
                      opt$D, profit=profit, penalty=L2(apples["L2"]), seed=seeds[x])),
  fixed = lapply(reps, function(x) simulate_optim(f, pars, x_grid, h_grid, x0, 
                         fixed_policy, z_g, z_m, z_i, 
                         opt$D, profit=profit, penalty=fixed(apples["fixed"]), seed=seeds[x]))

)
```

```{r tidy, dependson="simulate_policy"}
#Make data tidy (melt), fast (data.tables), and nicely labeled.
dat <- melt(sims, id=names(sims[[1]][[1]]))  
dt <- data.table(dat)
setnames(dt, "L2", "replicate") # names are nice
setnames(dt, "L1", "penalty_fn") # names are nice
```

```{r profit_calcs, dependson="tidy"}
# Profit when accounting for penalty when present
optimal_cost <- dt[, sum(profit_fishing - policy_cost), by=penalty_fn ]

# Profit when ignoring penalty when present
ignore_when_present <- dt[, sum(profit_fishing_alt - policy_cost_alt), by=penalty_fn]

# Profit when assuming penalty when it is absent
assume_when_absent <- dt[, sum(profit_fishing), by=penalty_fn]

# Profit when ignoring penalty when it is absent
optimal_free <- dt[, sum(profit_fishing_alt), by=penalty_fn]

# Normalize by the optimal 
ignore_fraction <- ignore_when_present$V1/optimal_free$V1 # common normalization
assume_fraction <- assume_when_absent$V1/optimal_free$V1
assume_when_absent <- cbind(assume_when_absent, assume_fraction = assume_fraction, normalize_optimal_free=optimal_free$V1, normalize_optimal_cost = optimal_cost$V1)
ignore_when_present <- cbind(ignore_when_present, ignore_fraction = ignore_fraction)

# Name and merge columns
setnames(ignore_when_present, "V1", "ignore_cost")
setnames(assume_when_absent, "V1", "assume_cost")
error_costs <- merge(ignore_when_present, assume_when_absent, "penalty_fn")

# print_npv                             # theoretically acheivable profits under these costs
# optimal_cost$V1/optimal_free$V1       # actually realized

error_costs <- cbind(error_costs, sigma_g = sigma_g, reduction = reduction)

```{r helper_fn_1}
fig4 <- function(fraction_lost){
closest <- function(x, v){
  which.min(abs(v-x))
}
dt_npv <- data.table(fees)
index <- dt_npv[,closest(fraction_lost, (npv0-value)/npv0), by=variable]
apples_index <- index$V1
names(apples_index) = index$variable
apples <- c2[index$V1]
names(apples) = index$variable

L2_policy <- policies$L2[[apples_index["L2"]]]$D
L1_policy <- policies$L1[[apples_index["L1"]]]$D
fixed_policy <- policies$fixed[[apples_index["fixed"]]]$D
free_increase_policy <- policies$free_increase[[apples_index["free_increase"]]]$D
free_decrease_policy <- policies$free_decrease[[apples_index["free_decrease"]]]$D
quad_policy <- policies$quad[[apples_index["quad"]]]$D

quad_profit <- profit_harvest(price = price, c0 = c0, c1 = apples["quad"]) 
sims <- lapply(1:50, function(reps) list(
  L1 = simulate_optim(f, pars, x_grid, h_grid, x0, 
                      L1_policy, z_g, z_m, z_i, 
                      opt$D, profit=profit, penalty=L1(apples["L1"])), 
  L2 = simulate_optim(f, pars, x_grid, h_grid, x0, 
                      L2_policy, z_g, z_m, z_i, 
                      opt$D, profit=profit, penalty=L2(apples["L2"])),
  fixed = simulate_optim(f, pars, x_grid, h_grid, x0, 
                         fixed_policy, z_g, z_m, z_i, 
                         opt$D, profit=profit, penalty=fixed(apples["fixed"]))
#  increase = simulate_optim(f, pars, x_grid, h_grid, x0, 
#                            free_increase_policy, z_g, z_m, z_i, 
#                            opt$D, profit=profit, penalty= free_increase(apples["increase"])),
#  decrease = simulate_optim(f, pars, x_grid, h_grid, x0, 
#                            free_decrease_policy, z_g, z_m, z_i, 
#                            opt$D, profit=profit, penalty= free_decrease(apples["decrease"])),
#  quad = simulate_optim(f, pars, x_grid, h_grid, x0, 
#                            quad_policy, z_g, z_m, z_i, 
#                            opt$D, profit=quad_profit, penalty= none)
  ))

  #Make data tidy (melt), fast (data.tables), and nicely labeled.
  dat <- melt(sims, id=names(sims[[1]][[1]]))  
  dt <- data.table(dat)
  setnames(dt, "L1", "reps") # names are nice
  setnames(dt, "L2", "penalty_fn") # names are nice

  dt
}
```

```{r sim_at_each_apple}
frac_lost <- seq(0,1, length=20)
sims_at_each_apple <- lapply(frac_lost, fig4)
```

```{r stats-defs}
summary_stats <- function(dt){
 harvest_var <- dt[,var(harvest), by=c("penalty_fn", "reps")]$V1
 harvest_acor <- dt[,acf(harvest, plot=F)$acf[2], by=c("penalty_fn", "reps")]$V1
 stock_var <- dt[,var(fishstock), by=c("penalty_fn", "reps")]$V1
 stock_acor <- dt[,acf(fishstock, plot=F)$acf[2], by=c("penalty_fn", "reps")]$V1
 xcor <- dt[, ccf(fishstock, harvest, plot=FALSE)$acf[2], by=c("penalty_fn", "reps")]
 setnames(xcor, "V1", "cross-correlation")

 df <- data.frame('stock.variance' = stock_var, 
                  'stock.autocorrelation' = stock_acor, 
                  'harvest.variance' = harvest_var, 
                  'harvest.autocorrelation' = harvest_acor, 
                  xcor)

 # long format
 melt(df, id=c("penalty_fn", "reps"))
 }
```

```{r summary-stat-calc, dependson="sim_at_each_apple"}
stats_summaries <- lapply(sims_at_each_apple, summary_stats)
# reformat list of data-frames as data frame with apple coef as a factor
stats_df <- melt(stats_summaries, id=names(stats_summaries[[1]]))
stats_df$L1 <- frac_lost[stats_df$L1]
names(stats_df) <- c("penalty_fn", "reps", "variable", "value", "penalty_fraction")
```

```{r split-factors}
# make statistic and timeseries separate factors
figure4_df <- stats_df
figure4_df$statistic <- figure4_df$variable
figure4_df$timeseries <- figure4_df$variable
stat_map <- c(harvest.variance = 'variance', stock.variance='variance', stock.autocorrelation='autocor',harvest.autocorrelation='autocor', cross.correlation='cross.correlation')
series_map <- c(harvest.variance = 'harvest', stock.variance='stock', stock.autocorrelation='stock',harvest.autocorrelation='harvest', cross.correlation='both')

figure4_df$statistic <- stat_map[figure4_df$statistic]
figure4_df$timeseries <- series_map[figure4_df$timeseries]
```

```{r errors-table}

source("components/compute_error_table.R")

reduction_list <- c(.1, .15, .2, .25, .3)
out <- lapply(reduction_list, function(r)
    compute_error_table(r = r))   
who <- names(out[[1]])
error_df <- melt(out, id=who)

````

<!-- ENDCODE -->

@@

```{r Figure_1, fig.cap="Expected net present value by functional form of penalty. Horizontal axis shows the coefficient $c_i$ governing the magnitude of the policy cost, while vertical axis shows fraction of the maximum expected net present value dissipated by the policy cost, $\\tfrac{NPV_0 - NPV_i}{NPV_0}$.  The horizontal line indicates a value of the stock that is reduced by `r reduction * 100`% from the maximum expected value in the absence of policy adjustment costs, $NPV_0({\\bf h_0^*})$. Selecting the coefficient $c_i$ corresponding to this value in each functional form allows us to make consistent comparisons across the different functional forms of policy costs.", dependson="quadcosts", fig.width=4, fig.height=3}

ggplot(fees, aes(c2, (npv0-value)/npv0, lty=variable)) + 
  geom_line() + 
  geom_hline(aes(hintercept=reduction), linetype=4) + 
  xlab("penalty coefficient") +
  ylab("Reduction in net present value") + 
  theme_tufte()

@@

ggplot(fig2_df, aes(time, value, col=baseline)) +
  geom_line() +
  facet_grid(variable~penalty_fn) + 
  labs(x="time", y="stock size", title = "Example Stock & Harvest Dynamics")  + 
  theme_tufte()

@@

ggplot(subset(figure4_df, statistic != "cross.correlation"), aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn))  +
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  facet_grid(statistic ~ timeseries, scale="free") + theme_bw()

@@

```{r Table1}
who <- c("penalty_fn", "ignore_fraction", "assume_fraction", "reduction")
table1 <- arrange(error_df[who], reduction) 
names(table1) = c("penalty.fn", "ignoring", "assuming", "reduction")
pandoc.table(table1, caption=caption)
```

```{r Table1-Figure, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, NPV0.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs."}
table1_long <- melt(table1, id = c('penalty.fn', 'reduction'))
table1_long$reduction <- as.factor(table1_long$reduction)
ggplot(table1_long, aes(penalty.fn, value, fill = variable)) + geom_bar(stat="identity", position="dodge") + facet_wrap(~ reduction) + theme_bw()
```

@@

ggplot(fig1s_df, aes(time, value, col=baseline, lty=replicate)) +
  geom_line() +
  facet_grid(variable~penalty_fn) + 
  labs(x="time", y="stock size", title = "Example Stock & Harvest Dynamics")  + 
  theme_tufte()

>>>>

output:
  pdf_document:
    keep_tex: true
    template: components/elsarticle.latex
    fig_caption: true
---

```{r dependencies, include=FALSE, cache=FALSE}
#source("components/install.R")
```
```{r analysis, include=FALSE}
source("components/analysis.R")
```

@@

```{r Figure_1, fig.cap="Expected net present value by functional form of penalty. Horizontal axis shows the coefficient $c_i$ governing the magnitude of the policy cost, while vertical axis shows fraction of the maximum expected net present value dissipated by the policy cost, $\\tfrac{NPV_0 - NPV_i}{NPV_0}$.  The horizontal line indicates a value of the stock that is reduced by `r reduction * 100`% from the maximum expected value in the absence of policy adjustment costs, $NPV_0({\\bf h_0^*})$. Selecting the coefficient $c_i$ corresponding to this value in each functional form allows us to make consistent comparisons across the different functional forms of policy costs.", dependson="quadcosts", fig.width=4, fig.height=3, cache=FALSE}
ggplot(fees, aes(c2, (npv0-value)/npv0, lty=variable)) + 
  geom_line() + 
  geom_hline(aes(yintercept=reduction), linetype=4) + 
  xlab("penalty coefficient") +
  ylab("Reduction in net present value")

@@

ggplot(fig2_df, aes(time, value, col=baseline)) +
  geom_line() +
  facet_grid(variable~penalty_fn) + 
  labs(x="time", y="stock size", title = "Example Stock & Harvest Dynamics")

@@

ggplot(subset(figure4_df, statistic != "cross.correlation"), aes(penalty_fraction, value, fill=penalty_fn, col=penalty_fn))  +
  stat_summary(fun.y = mean, 
               fun.ymin = function(x) mean(x) - sd(x), 
               fun.ymax = function(x) mean(x) + sd(x), 
               geom = "ribbon", alpha=0.3, colour=NA) +
  stat_summary(fun.y = mean, geom = "line") + 
  facet_grid(statistic ~ timeseries, scale="free_y")

@@

```{r Table1, results="asis"}
who <- c("penalty_fn", "ignore_fraction", "assume_fraction", "reduction")
table1 <- arrange(error_df[who], reduction) 
names(table1) = c("penalty.fn", "ignoring", "assuming", "reduction")
pandoc.table(table1, caption=caption)
```

```{r Table1-Figure, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, NPV0.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs."}
table1_long <- melt(table1, id = c('penalty.fn', 'reduction'))
table1_long$reduction <- as.factor(table1_long$reduction)
ggplot(table1_long, aes(penalty.fn, value, fill = variable)) + geom_bar(stat="identity", position="dodge") + facet_wrap(~reduction) 
```

@@

ggplot(fig1s_df, aes(time, value, col=baseline, lty=replicate)) +
  geom_line() +
  facet_grid(variable~penalty_fn) + 
  labs(x="time", y="stock size", title = "Example Stock & Harvest Dynamics")
